---
title: "TidyR_homework"
author: "YOUR NAME HERE"
date: "2026-03-31"
output: pdf_document
---

## TidyR Homework

In this assignment, you will be asked to select a dataset of your choice, identify any 'untidy' characteristics, make it tidy, and do some exploratory analysis. 

## Question 1 (1 point)

Select a dataset from https://github.com/rfordatascience/tidytuesday/tree/main/data and import it into RStudio. 

If your dataset references a paper, like the one we did in class, you can download it from its data repository. If it is only found on the tidy tuesday github, you may need to use one of these functions:
tidytuesdayR::tt_load 
readr::read_csv

Show me the first *10 rows* of your data (1 point).

```{r Question 1}
#load a dataset

package_list <- c("dplyr", "readr", "tidyverse", "tidyr", "lubridate", "dedupewider")

for (pckg in package_list) {
  if (!requireNamespace(pckg, quietly = TRUE)) {
    install.packages(pckg, dependencies = TRUE)
  }
  library(pckg, character.only = TRUE)
}

dls_protein_corona_size = read_csv("20250502.csv")

#output the first 10 rows of your data

head(dls_protein_corona_size, 10)

```

## Question 2 (3 points)

Identify at least 3 non-tidy characteristics of your data or things you would like to change in this dataset (1 point ea):

* 1. Each set of sample measurement duplicates is preceded by a header row, so that the column headers appear multiple times in each column.

* 2. The size values (in nm) and the intensity values corresponding to those sizes are contained in the same row.

* 3. The concentration, elemental composition, and time information is all contained in the sample name.

* 4. The "S/W Version" and "Serial Number" are useless information because they are the same for all samples.

* 5. The measurement date and time contains characters, but I would like to convert it to numeric.

* 6. The volumes data .

Show any code you used to identify these characteristics:

```{r Q2}

```

## Question 3 (5 points)

Fix each of your issues. Use comments to show where each of your issues listed above was fixed. Use pipes to connect fixing all of your issues together!

```{r Q3}

dls_protein_corona_size_cleaned = dls_protein_corona_size %>%
  filter(Type == "Size",) %>% #Removes interdigitated header rows, addresses issue 1 above
  select(-c(`S/W Version`, `Serial Number`)) %>% #Removes "S/W Version" and "Serial Number" columns, addresses issue 4 above
  mutate(
    datetime = as.POSIXct(`Measurement Date and Time`,
                          format = "%A, %B %d, %Y %I:%M:%S %p"),
    measurement_time_numeric = hour(datetime) + minute(datetime)/60 + second(datetime)/3600) %>% #Converts "Measurement Date and Time" column to numeric decimal hours, addresses issue 5 above
  select(-c(`Measurement Date and Time`)) #Removes "Measurement Date and Time" column

sizes_and_intensities = dls_protein_corona_size_cleaned %>% 
  select(c(`Sample Name`, starts_with("Sizes"), starts_with("Intensities"))) #Creates a new dataframe specifically for intensity versus size data

colnames(sizes_and_intensities)[72:141] = sizes_and_intensities[1, 2:71] #Renames column headers of intensity data with size values from row 1, addresses issue 2 above

sizes_and_intensities_cleaned = sizes_and_intensities %>%
  select(-c(starts_with("Sizes"))) %>% #Removes repetitive size data
  pivot_longer(cols = -1, names_to = "size_nm", values_to = "intensity_percent") #Pivots table so that size values and intensities are in separate columns

sizes_and_intensities_cleaned$size_nm = as.numeric(sizes_and_intensities_cleaned$size_nm)
sizes_and_intensities_cleaned$intensity_percent = as.numeric(sizes_and_intensities_cleaned$intensity_percent)

dls_protein_corona_size_concentration = dls_protein_corona_size_cleaned %>%
  select(-c(starts_with("Intensities"), starts_with("Sizes"), starts_with("Volumes"))) %>%
  mutate(concentration = str_extract(`Sample Name`, "(?<=_)\\d+\\.?\\d*(?=mg/mL)") %>% as.factor()) %>%
  drop_na(`concentration`) %>%
  mutate(elapsed_time = `measurement_time_numeric` - `measurement_time_numeric`[1])

```

## Question 4 (2 points)

List a question you would like to answer using this dataset:

How does the average hydrodynamic particle diameter change as a function of incubation time and as a function of protein concentration?

## Question 5 (5 points)

Using stats, plots, or both, answer your question. Make sure you output any stats or plots you create in this code chunk (i.e., that it shows up in the knitted document). 

NOTE: you do not have to use a formal stats test if you are not familiar! Feel free to just use summary stats and/or figures to answer your question. 

```{r Q5}

sizes_and_intensities_cleaned %>%
  #filter(`Sample Name` == "SeNP_only_15min 1") %>%
  ggplot(aes(x=log(size_nm), y=intensity_percent)) +
  geom_point() +
  facet_wrap(~ `Sample Name`)

dls_protein_corona_size_concentration %>%
  ggplot(aes(x=elapsed_time, y=`Z-Average (d.nm)`, color=concentration)) +
  geom_point()
```

Provide the answer (in words) to your question below:


## Question 6 (4 points)

Find two functions we haven't introduced you to that are a part of tidyR, dplyr, or other related tidyverse packages. Provide the function name (including the package it's from) and what it does.




